%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Master file SED-ML specification 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\listfiles
\documentclass[pdftex,rgb,dvipsnames,svgnames,hyperref,table]{report}
\usepackage{tocvsec2}
\usepackage{latexsym}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{scalefnt}
\usepackage{suffix}
%\usepackage{textcmds}

% layout (initialy done for the SBGN project)
\input{sources/styles/latex-style}

% packages and commands
\input{sources/styles/latex-macros}
\input{sbmlpkgspec}

\begin{document}
\input{sources/frontpage}
\sedmltableofcontents
\newpage
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%% INTRODUCTION
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\input{sources/introduction}

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%% TECHNICAL SPECIFICATION
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\chapter{SED-ML technical specification}
\label{chp:specification}
This document is the technical specification of SED-ML \currentLV. The corresponding UML class diagrams \changed{are included throughout this document.}. Please note, the \changed{UML diagrams} do not capture all concepts of SED-ML. Example simulation experiments encoded with SED-ML are provided in Appendix~\ref{app:examples}. Complete examples with model files are available at \url{http://sed-ml.org/}.
%The XML Schema is provided in Appendix~\ref{app:schema}.

%\sedfig[width=\textwidth]{images/uml/all}{The SED-ML \currentLV UML class diagram}{fig:sedML}

\settocdepth{subsubsection}

%% ~~~ GENERAL LANGUAGE ELEMENTS ~~~
\input{sources/generalElements}

%% ~~~ COMPONENTS ~~~
\pagebreak
\section{SED-ML Components}
\label{sec:components}
This section describes the major components of SED-ML. \changed{Each subsection includes UML diagrams for the relevant classes.}  Example simulation experiments are provided in Appendix~\ref{app:examples}. Complete examples with model files are available at \url{http://sed-ml.org/}.%, and the XML Schema is listed in Appendix~\ref{app:schema}.

%% SEDML
\input{sources/sedml}
%% DATA
\input{sources/data}
%% MODEL
\input{sources/model}
%% SIMULATION
\input{sources/simulation}
%% TASK
\input{sources/task}
%% DATA GENERATOR
\input{sources/dataGenerator}
%% OUTPUT
\input{sources/output}
%% STYLES
\input{sources/style}

%% ~~~ CONCEPTS ~~~
\input{sources/concepts}

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%% ACKNOWLEDGMENTS
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\chapter{Acknowledgements}
\label{chp:acknowledgments}
The SED-ML specification is developed with the input of many people. The following individuals have contributed to the SED-ML specifications.

\begin{itemize}
\item Richard Adams (Editor, 2011-2012)
\item Frank Bergmann (Editor, 2011-2014, 2020-2022)
\item Jonathan Cooper (Editor, 2012-2015)
\item Alan Garny (Editor, 2018-2020)
\item Tomas Helikar (Editor, 2021-2023)
\item Jonathan Karr (Editor, 2021-2023)
\item Matthias K\"{o}nig (Editor, 2017-2019, 2020-2022)
\item David Nickerson (Editor, 2011-2013, 2015-2017, 2019-2021)
\item Nicolas Le Nov{\`e}re (editorial advisor, 2011-2012, 2013)
\item Brett Olivier (Editor, 2015-2017)
\item Andrew Miller (Editor, 2011-2012)
\item Ion Moraru (Editor, 2014-2016)
\item Sven Sahle (Editor, 2014-2016)
\item Herbert Sauro (Editor, 2018-2020)
\item Lucian Smith (Editor, 2016-2018)
\item Dagmar Waltemath (Editor, 2011-2014, 2017-2019)
\end{itemize}

In addition, we thank the many members of the SED-ML coummunity who have contributed to the development of SED-ML.

\appendix

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% EXAMPLES
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\chapter{Examples}
\label{app:examples}
This appendix presents several SED-ML examples. Complete versions of these and additional examples are available as \hyperref[sec:archive]{Combine archives} at \url{http://sed-ml.org/}. These examples illustrate the main features of SED-ML. Please note, these examples do not demonstrate the full capabilities of SED-ML. The specifications of SED-ML (Chapter~\ref{chp:specification}) provide a more comprehensive view of the simulation experiments that can be captured with SED-ML. 

The examples presented here involved models encoded in CellML and SBML. \changed{Please note, SED-ML can be used with additional model languages. See} Section~\ref{sec:languageURI} for more information about using SED-ML with additional model languages. Example SED-ML files for additional model languages are available at \url{https://run.biosimulations.org/}.

\input{sources/examples}

\chapter{Validation}
\label{app:validation}
\begin{blockChanged}
\input{sources/apdx-validation}
\end{blockChanged}

%% ~~~ VALIDATION ~~~
%\input{sources/apdx-validation}



% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% XML SCHEMA
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%\chapter{XML Schema}
%\label{app:schema}
%Listing \ref{lst:schema} shows the full SED-ML XML Schema.
%\myXmlImport{The SED-ML XML Schema definition}{lst:schema}{schema/sed-ml-L1-V4.xsd}

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% REFERENCES
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\bibliographystyle{abbrv}
\bibliography{sed-ml}
\end{document}
