\subsubsection{\element{target}}
\label{sec:target}
An instance of \concept{Variable} can refer to a model constituent inside a particular \hyperref[class:model]{model} through an \concept{XPath} expression stored in the \concept{target} attribute. 
%
XPath can be used to unambiguously identify an element or attribute in an XML file.

The \concept{target} attribute may also be used to reference an entity within the SED-ML file itself, by containing a fragment identifier consisting of a hash character (\code{\#}) followed by the \concept{id} of the desired element.
As of SED-ML \LoneVtwo this is only used to refer to \hyperref[sec:ranges]{ranges} within a \hyperref[class:repeatedTask]{repeatedTask} (see Listing~\ref{lst:repeatedTask} for an example).


\lsttexta{target}{target}
%
\begin{myXmlLst}{SED-ML \code{target} definition}{lst:target}
  <listOfVariables>
    <variable id="v1" name="TetR protein" taskReference="task1" 
      target="/sbml:sbml/sbml:listOfSpecies/sbml:species[@id='PY']" />
  </listOfVariables>
\end{myXmlLst}
%
It should be noted that the identifier and names inside the SED-ML document do not \emph{have} to match the identifiers and names that the model and its constituents carry in the model definition.
In listing \vref{lst:target}, the variable with ID \code{v1} is defined. It is described as the \code{TetR protein}. The reference points to a species in the referenced SBML model. The particular species can be identified through its ID in the SBML model, namely \code{PY}.
However, SED-ML also permits using identical identifiers and names as in the referenced models. The following Listing~\vref{lst:sedmlVariable} is another valid example for the specification of a variable, but uses the same naming in the variable definition as in the original model (as opposed to Listing~\ref{lst:target}):
%
\begin{myXmlLst}{SED-ML variable definition using the original model identifier and name in SED-ML}{lst:sedmlVariable}
   <listOfVariables>
    <variable id="PY" name="TetR protein"  taskReference="task1" 
     target="/sbml:sbml/sbml:listOfSpecies/sbml:species[@id='PY']" />
   </listOfVariables>
\end{myXmlLst}
%

%
\begin{myXmlLst}{Species definition in the referenced model (extracted from \url{urn:miriam:biomodels.db:BIOMD0000000012})}{lst:sbmlModel}
<sbml [..]>
 <listOfSpecies]
  <species metaid="PY" id="PY" name="TetR protein" [..]>
   [..]
  </species>
 </listOfSpecies>
 [..]
</sbml>
\end{myXmlLst}
%

The XPath expression used in the \concept{\code{target}} attribute unambiguously leads to the particular place in the XML SBML model -- the species is to be found in the \emph{sbml} element, and there inside the \emph{listOfSpecies} (Listing~\vref{lst:sbmlModel}).
Note that while it is possible to write XPath expressions that select multiple nodes within a referenced model, when used within a \concept{target} attribute a single element or attribute \emph{must} be selected by the expression.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../sed-ml-L1V2"
%%% End: 
