\documentclass{article}


\newcommand{\code}[1]{\texttt{#1}}

\begin{document}
\title{Changes in the specification document for SED-ML Level 1 Version 2}
\author{The SED-ML Editors}

\maketitle

The following is a list of changes applied to the specification document of SED-ML L1V1 to transform it into L1V2. The changes were made based on the votes for the algorithm/KiSAO proposal and the Nested proposel.

\section{To dos}
\emph{This is a list of things that still need to be done. Please provide the file name with each change.}

\begin{enumerate}
\item Update the section on SED-ML archive
\end{enumerate}

\subsection{Figures}
\begin{enumerate}
\item Update UML diagram 2.29 (The Algorithm class) - add algorithmParameter (section 2.4.3.2)
\item Add UML diagram for algorithmParameter class (section 2.4.3.3) 
\item Add figure for listOfAlgorithmParameters (section 2.3.8.6)
\end{enumerate}

\subsection{Example code}
\begin{enumerate}
\item Check all simulation examples to include algorithm parameters
\end{enumerate}
\subsection{Textual description}

\subsection{XML schema}
\begin{enumerate}
\item Add listOfAlgorithmParameters/algorithmParameter to schema
\end{enumerate}
% % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Applied changes}
\emph{The following is a listing of all changes that have been made in the specification document for Level 1 Version 2. Please provide the file name with each change.}

\subsection{Figures}

\subsection{Example code}
\begin{enumerate}
\item example added for algorithmParameter element (section 2.4.3.3)
\item example added for listOfAlgorithmParameter element (section 2.3.8.6)
\end{enumerate}

\subsection{Textual description}

\begin{enumerate}
\item Mention of algorithmParameter class in algorithm class, section 2.4.3.2: "The Algorithm class contains an optional list of parameters (algorithmParameter)that can be set for the particular algorithm used in the simulation."
\item New class: AlgorithmParameter "The AlgorithmParameter class allows to set tolerances and other parameters for a particular simulation
algorithm. The set of possible parameters is determined by the algorithm that is referenced by the
kisaoID in the Algorithm class. Parameters of simulation algorithms are unambiguously referenced by
the mandatory kisaoID attribute. Their value is set in the mandatory value attribute."
\item New list element: listOfAlgorithmParameters (section 2.3.8.6)
\end{enumerate}

\subsection{tex}

\begin{enumerate}
\item Changed tex-master from  \code{TeX-master: "../sed-ml-L1V1"} to  \code{TeX-master: "../sed-ml-L1V2"}
\end{enumerate}
\end{document}
