\label{class:notes}

A \concept{note} is considered a  human-readable description of the element it is assigned to. It serves to display information to the user. 
\concept{Notes} contains any valid XHTML \citep{P+02}, ranging from short comments to whole HTML pages for display in a Web browser. 

\concept{Notes} has a mandatory attribute \hyperref[sec:xmlns]{xmlns} to declare the \code{XHTML} namespace. It does not have any further subclasses nor attributes associated to it (see \tab{notes}).

\tabtext{notes}{Notes}

%
\begin{table}[ht]
\center
\begin{tabular}{|l|l|}
\hline
\textbf{\attribute} & \textbf{\desc}\\
\hline
xmlns & \refpage{sec:xmlns} \\
\hline
\hline
\textbf{\subelements} & \textbf{\desc}\\
\hline
\emph{none} & \\
\hline
\end{tabular}
\label{tab:notes}
\caption{\tabcap{Notes}}
\end{table}

The namespace URL for \code{XHTML} content inside the \hyperref[class:notes]{Notes} class is \url{http://www.w3.org/1999/xhtml}. It may either be declared in the \hyperref[class:sed-ml]{\code{sedML} XML element}, or directly for use in each \code{notes} element of the XML file. For further options of how to set the namespace and detailed examples, please refer to (\citep{HBH+10}, p. 14).

Listing \ref{lst:notes} shows the use of the \element{notes} element in a SED-ML file. 
\begin{myXmlLst}{The \element{notes} element}{lst:notes}
<sedML version="0.1">
 <notes http://www.w3.org/1999/xhtml>
  The enclosed simulation description shows the oscillating behaviour of the Repressilator model using deterministic and stochastic simulators.
 </notes>
</sedML>
\end{myXmlLst}
In this example the namespace declaration is inside the \element{notes} element and the note is related to the \element{sedML} root element of the SED-ML file. A note may, however, occur inside \emph{any} SED-ML XML element, except \element{note} itself and \hyperref[class:annotation]{\code{annotation}}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../sed-ml-L1V1"
%%% End: 
